Intrinsic alignments from projected tidal fields: non-linear impact on cosmic shear probes

### 2024-03-18 paper_IA.tex
Added preliminary HOD results

Cosmology, gravitational lensing, cosmic shear, 2pt, beyond-2pt, simulations for signal and for secondary signal/systematics, IA-infusion, this paper...

 %----------
\section{Cosmic shear}
\label{sec:theory}
%----------

As mentioned in the introduction, cosmic shear can be measured and modelled from a variety of methods, many of which are used in this paper and presented in this section.  
%-------------------------
\subsection{$\gamma$-2PCF}
\label{subsec:wl-th}

%---------------
\begin{figure}
\includegraphics[width=\columnwidth]{graphs/Nz}
\caption{Redshift distribution in our simulations, taken from the Year-1 specifications of the LSST Science Requirement Document \citep{LSST-SRD} {[\it check Y1 vs Y10}].}
\label{fig:Nz}
\end{figure}
%--------------

 Shear two-point correlation functions ($\gamma$-2PCFs hereafter) can be predicted with percent-level precision and are therefore an ideal quantity to validate weak lensing simulations. In the Limber approximation\footnote{See \citet{Kilbinger17} for a comparison between the Limber approximation and the exact calculations.}, the lensing power spectrum $C^{ij}_{\ell}$ is obtained from an integral over the three-dimensional matter power spectrum $P_{\delta}(k,z)$:
  \begin{eqnarray}
C_{\ell}^{ij} = \int_0^{\chi_{\rm H}}  \frac{q^i(\chi) \,q^j(\chi)}{\chi^2} \, P_{\delta}\, \bigg(\frac{\ell+1/2}{\chi},z(\chi)\bigg) \ {\rm d}\chi,
\label{eq:C_ell}
\end{eqnarray}
where $\chi_{\rm H}$ is the co-moving distance to the horizon, $c$ is the speed of light, $H_0$ the Hubble parameter,  and the  lensing kernels $q^{i}$ are given by:
\begin{eqnarray}
q^i(\chi) = \frac{3}{2}\Omega_{\rm m} \, \bigg(\frac{H_0}{c} \bigg)^2 \frac{\chi}{a(\chi)} \int_{\chi}^{\chi_{\rm H}} n^i(\chi')\frac{{\rm d}z}{{\rm d} \chi'}\frac{\chi' - \chi}{\chi'}{\rm d}\chi' \, .
\label{eq:q_lensing}
\end{eqnarray}
In the previous expression, $n^i(z)$ refers to the redshift distribution  in tomographic bin `$i$', while $a(\chi)$ is the scale factor at comoving distance $\chi$ from the observer {[\it might need to define $\Omega_m$ here]}.  
 The matter power spectrum is computed from  {\sc Halofit} \citep{Takahashi2012}, however  other public tools provide $P_{\delta}(k,z)$ predictions, including e.g. {\sc HMcode} \citep{HMCode2020}, the {\sc EuclidEmulator} \citep{EuclidEmulator}, the {\sc BACCO} emulator \citep{BACCOEmulator} or the Mira Titan emulator \citep{miraTitan}. 
 Predictions for the $\gamma$-2PCFs are finally computed from Eq. (\ref{eq:C_ell}) as:
 \begin{eqnarray}
\xi_{\pm}^{ij}(\vartheta) = \frac{1}{2\pi}\int_0^{\infty} C_{\ell}^{ij} \, J_{0/4}(\ell \vartheta) \, \ell \, {\rm d}\ell,
\label{eq:xipm_th}
\end{eqnarray}
where $J_{0/4}(x)$ are Bessel functions of the first kind. In this paper, these calculations are carried out by {\sc pyccl}\footnote{{\sc pyccl}:pypi.org/project/pyccl}  \citep{CCL} {\it or do we use cosmosis in the end?]}.
The $n^i(z)$ are taken from the LSST Science Requirement Document \citep{LSST-SRD} and given by
 \begin{eqnarray}
n(z) = N_0 z^2 {\rm exp} \left[-\left(\frac{z}{z_0}\right)^\alpha \right]\, ,
\label{eq:nz}
\end{eqnarray}
with $(z_0, \alpha) = (0.13, 0.78)$ and $N_0$ normalised to provide a number density of $n_{\rm eff}$ = 3.0 gal arcmin$^{-2}$. This lower than the expected number density for the first data release ($n_{\rm eff}\sim10$ gal arcmin$^{-2}$), but is large enough to validate our methods, and makes the calculation more tractable. The $n(z)$ is further split into five equi-populated tomographic bins, and smoothed with a Gaussian filter of width $\sigma = 0.05(1+z)$ in order to mimic the photometric selection process [{\it confirm with Niko}].  The resulting distributions\footnote{The construction of these tomographic $n(z)$ can be reproduced in {github...CCL\_SRD\_Niko}.} are shown in Figure \ref{fig:Nz}.

The cosmic shear signal can be measured from the ellipticities $\epsilon_{1/2}$ of simulated or observed galaxies, which, in absence of systematics and secondary signals, are unbiased estimators  of the cosmic shear components $\gamma_{1/2}$. The $\gamma$-2PCF are then estimated as:
 \begin{eqnarray}
 \widehat{\xi_{\pm}^{ij}}(\vartheta) = \frac{\sum_{a,b} w_a w_b \left[\epsilon_{a, \rm t}^i\epsilon_{b, \rm t}^j     \pm \epsilon_{a, \times}^i\epsilon_{b, \times}^j    \right]\Delta_{ab}(\vartheta)}{\sum_{a,b} w_a w_b} \, ,
 \label{eq:2PCF_estimator}
  \end{eqnarray}
where the sum is over all pairs of galaxies $a,b$ separated by an angular separation $\vartheta$, taken from tomographic bins $i,j$. The tangential/cross components of their ellipticities are denoted as $\epsilon_{\rm t/\times}$, respectively. 
$\Delta_{ab}(\vartheta)$ is the binning operator, equal to unity if the angular separation between the two galaxies falls within the $\vartheta$-bin, and zero otherwise, and the shape weights $w_{a,b}$ are set to unity. In this paper, we construct lensing catalogues from numerical simulations (described in Sec. \ref{sec:sims}), from which we measure $ \widehat{\xi_{\pm}^{ij}}(\vartheta)$ with {\sc Treecorr} \citep{TreeCorr}. We set therein the {\sc bin\_slop} parameter to 0.05, then compute the correlations in nine logarithmically-spaced angular bins ranging from 0.5 to 475.5 arcmin.

 %-------------------------
\subsection{Beyond-2pt}
\label{subsec:beyond-2pt}

Non-Gaussian statistics, often referred to `beyond-2pt' statistics, have the potential to extract cosmological information stored in the complex phases of the density field, whereas two-point functions such as the $\gamma$-2PCF can only access that encoded in the complex amplitudes. No optimal estimator has been identified to-date as being able to capture `all' information, but the numerous studies presented in the introduction all show a clear gain in constraining power, especially when used in combination with 2pt functions. Some of these take as input the galaxy catalogues directly, while others require instead convergence maps. Therefore, we produce convergence maps from the catalogues by assigning the shear signal onto spherical {\sc Healpix}\footnote{{\sc Healpix}: http://healpix.sf.net}  maps \citep{healpix} with $N_{\rm side}=4096$, then using the standard Kaiser \& Squires inversion technique \citep{KaiserSquires} implemented in {\sc healpy} to reconstruct a convergence maps with the same angular resolution, including a baseline 2.0 arcmin smoothing beam in the inverse transform. Note that this smoothing scales is quite small; some probes are difficult to model in that case, and would instead prefer larger smoothing scales. We take this into account on a case-by-case basis, as described below.

  {\it [Discuss tomography...]}

\subsubsection{Catalogue-based estimators}

We consider the following shear-based estimators:
\begin{enumerate}
\item The squeezed bispectrum, has theoretical model (Anik, can you summarise the measurements and modelling, and provide references?)
\end{enumerate}

\subsubsection{Convergence-based estimators}

\begin{enumerate}
\item Convergence three-point correlation function: (Alejandro, can you summarise the measurements and modelling, and provide references?)
\item Peaks and Minima: local maxima and minima in the {\sc Healpix} convergence maps are counted in bins of signal-to-noise ratios, $\nu$. This is one of the most widely used non-Gaussian lensing statistics as it is simple to measure. Theoretical models have been developed to describe the largest peaks \citep[see][]{Shan18, HSCY1_Peaks_th}, however we report here the results on a much wider range of statistics, which can only be accurately modelled from numerical simulations, as in \citet{HD21, DESY3_Zuercher, HSCY1_Peaks_Sims, HD23}.
\item Lensing PDF, has theoretical model (Cora/Lina, can you summarise the measurements and modelling, and provide references?)
\item Map2 and Map3, has theoretical model  (Laila, can you summarise the measurements and modelling, and provide references?)
\item Minkowski functionals (Nisha, can you summarise the measurements and modelling, and provide references?)
\end{enumerate}

@@

Galaxies interact with the tidal forces caused by the large-scale structure  they live in, causing their intrinsic shapes to acquire correlated alignments that has nothing to do with the correlations caused by weak lensing. This therefore acts as a secondary signal that contaminates the shape correlation measurements carried out in a cosmic shear analysis. There is no consensus on the actual physical model that describes the IA signal, and even when adopting these, the free parameters they contain are only weakly constrained from the data \citep[see][for reviews]{IA_review1, IA_review2}.  %The most widely used model in the literature is the non-linear alignment model of \citet{NLA}, however it is now recognised that this is an effective model with limited precision, and the community is now progressively shifting towards more complex models. 
In this paper we consider four such IA models that we describe in the following sub-sections. In all cases, our models couple the galaxy intrinsic shapes with the local over-density and projected tidal field , then prescribe an intrinsic ellipticity tensor $\gamma_{ij}^{\rm IA}$, from which observed ellipticities are extracted as:
 \begin{eqnarray}
\epsilon_{1}^{\rm IA} = \gamma_{xx}^{\rm IA} - \gamma_{yy}^{\rm IA} \, , \hspace{1.5cm}\epsilon_{2}^{\rm IA} = 2 \gamma_{xy}^{\rm IA}\, .
\label{eq:tidal_th_deltaNLA}
\end{eqnarray}

@@

%The observed ellipticity of a galaxy ${\boldsymbol \epsilon}_{\rm obs} $ is a combination of its intrinsic shape ${\boldsymbol \epsilon}_{\rm int}$ and a cosmic shear signal ${\boldsymbol \gamma}$, the former of which can be further divided in a random component ${\boldsymbol \epsilon}^{\rm ran}$ and an alignment term ${\boldsymbol \epsilon}^{\rm IA}$.
The NLA model of \citet{NLA} is the most widely used IA model in the cosmic shear literature thus far. According to the NLA, IA are caused by a linear coupling between galaxy shapes and the non-linear large-scale tidal field at the galaxy position. In the context of  two-point functions, these intrinsic shapes contribute to an intrinsic-intrinsic ($II$) term and an intrinsic-shear coupling ($GI$) term \citep{Hirata2004}, both secondary signals to the true cosmic shear ($GG$) term, with the latter typically dominating the IA sector in cross-tomographic measurements.  The $II$ and $GI$ terms  can be both computed from the matter power spectrum as: 
 \begin{eqnarray}
P_{II}(k,z) =  \left(\frac{A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)}\right)^2a^4(z) P_{\delta}(k,z)
\label{eq:Pk_II_th}
\end{eqnarray}
and
 \begin{eqnarray}
P_{GI}(k,z) = - \frac{A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)}a^2(z) P_{\delta}(k,z) \, ,
\label{eq:Pk_GI_th}
\end{eqnarray}
which can then be past to the Limber integration (Eqs. \ref{eq:C_ell} and \ref{eq:xipm_th}) to compute  the secondary signals $\xi_{\pm}^{II}(\vartheta)$ and $\xi_{\pm}^{GI}(\vartheta)$. 
In the above expression,  $D(z)$ is the linear growth factor, $\bar{\rho}(z)$ is the mean matter density at redshift $z$ and $\bar{C_1}=5\times 10^{-14} M_{\odot}^{-1} h^{-2} {\rm Mpc}^3$ is a constant calibrated in \citet{Brown2002}. 
The strength of the tidal coupling is modulated by the amplitude parameter $A_{\rm IA}$, which is the main NLA parameter constrained by current cosmic shear surveys. Note that this model can be augmented by redshift and luminosity dependences, however we do not use these here.

The NLA intrinsic ellipticities $\epsilon_{1,2}^{\rm NLA}$ are related to tidal field $s_{ij}$ by:
 \begin{eqnarray}
\epsilon_1^{\rm NLA} = - \frac{A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)} (s_{xx} - s_{yy}) \,\,\,\,,   \epsilon_2^{\rm NLA} = - \frac{2 A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)} s_{xy}\;,
\label{eq:tidal_th}
\end{eqnarray}
where $s_{ij} = \partial_{ij}\phi$ are the Cartesian components of the tidal tensor of the gravitational potential. Note that the term `non-linear' in the model name can be  misleading, as it refers to the non-linear matter power spectrum $P(k)$ that is used in its calculations; the coupling between the intrinsic galaxy shapes and the tidal field is still linear.   Eq.( \ref{eq:tidal_th}) is used to assign intrinsic ellipticities to galaxies, given maps of the tidal field components $s_{xx}$, $s_{yy}$ and $s_{xy}$ (see Sec. \ref {subsec:IA_infusion}).

@@

As shown in \citet{Blazek2019}, one-loop perturbative calculations include another term, by which galaxies acquire an intrinsic alignment via a coupling between their angular momentum and the tidal field, which can be re-expressed as a quadratic coupling between the tidal field and the shape. In this tidal torque model (TT), intrinsic ellipticities are given by:
\begin{eqnarray}
\gamma_{ij}^{\rm IA, TT} = C_2 \left[ \sum_{k=x,y,z} s_{ik} s_{kj} -\frac{1}{3} \delta_{ij} s^2 \right] \, ,
\label{eq:tidal_th_TT}
\end{eqnarray}
where 
\begin{eqnarray}
C_2 = \frac{5 A_2 \bar{C_1} \Omega_{\rm m} \rho_{\rm crit}}{D^2(z)} \,  = \left[ \frac{-5 A_2}{A_1 D(z)} \right] C_1.
\end{eqnarray}
Under the approximation that line-of-sight alignments are mostly suppressed from cosmic shear measurement due to the broad lensing kernels, 
we show in  Appendix \ref{app:2d_TT} that the terms inside the square bracket in Eq. (\ref{eq:tidal_th_TT}) leads to:
\begin{eqnarray}
\epsilon_1^{\rm TT} = C_2  \left[ s_{xx}^2 - s_{yy}^2\right] \, , \epsilon_2^{\rm TT} = 2 C_2 s_{xy}\left[s_{xx}+s_{yy}  \right]\, ,
\end{eqnarray}
both quadratic in the tidal field components.

@@

@@

JHD acknowledges support from an STFC Ernest Rutherford Fellowship (project reference ST/S004858/1). Ray-tracing computations were carried on the NERSC [acknowledgments] ... HACC [acknowledgements] ... OuterRim [Acknowledgments]
\\

@@

%The SLICS numerical simulations can be found at http://slics.roe.ac.uk/, while the cosmo-SLICS can be made available upon request.

>>>>

Recent cosmic shear measurements from the Kilo Degree Survey\footnote{KiDS:kids.strw.leidenuniv.nl}, the Dark Energy Survey\footnote{DES:www.darkenergysurvey.org} and the Hyper Suprime Camera Survey\footnote{HSC:www.naoj.org/Projects/HSC} have established weak gravitational lensing as a competitive probe of cosmology \citep[see. e.g.,][]{KiDS1000_Asgari, KiDS1000_vdB, KiDS1000_Li, DESY3_Secco, DESY3_Amon, HSCY3_Cl, HSCY3_2PCF}, achieving  percent-level measurements of the structure growth parameter $S_8\equiv\sigma_8\sqrt{\Omega_{\rm m}/0.3}$. The parameters $\Omega_{\rm m}$ and $\sigma_8$, which respectively describe the abundance and the fluctuations of the matter density fluctuations  on scales of $8h^{-1}$ Mpc, are highly degenerate in the lensing signal, and additional date (e.g. galaxy clustering data) are required to measure the two individually \citep{Heymans, DES3x2pt, HSC3x2}. Despite these highly successful achievements, the current precision of cosmic shear cosmology is mainly limited by the large uncertainty in the intrinsic alignment  (IA) of galaxies, a secondary signal that tends to cancel some of the shape correlations produced by lensing \citep[see. e.g. ][for reviews on IA]{Joachim_IA, Kiessling_IA}. If unaccounted for, IA can bias by XYZ$\sigma$ the inferred cosmological parameters \citep{BlazekKrause}. Furthermore, a wrong IA model can also cause significant damage to the inference, as shown in \citet{DESY3_Secco} in the context of DES-Y3 analysis, or in \citep{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.

Different physical model have been developed to describe the origin and impact of the IA, including the linear non-linear alignment model \citep[][NLA hereafter]{BridleKing07}, the density-weighted NLA (aka $\delta$-NLA or extended-NLA) and the tidal torque model \citep[][TT hereafter]{Blazek2017}, which respectively assume a linear and quadratic coupling between galaxy ellipticities and the local tidal field. The halo model can also describe the IA signal as a function of halo properties such as their mass, concentration and shapes \citep{FortunaIA}. A number of observations have sought to place constraints on the parameters from these models \citep[e.g.][]{Singh, Sammuroff, Sammuroff, Christos, Fortuna, Johnston}, which seem to converge towards a strong colour-dependence (red bright galaxies are generally strongly aligned, with hints of a radial dependence), while blue galaxies are consistent with no-alignment. The selection effects play a crucial role, making it difficult to generalise these measurements to a different galaxy sample, therefore leaving behind a large uncertainty on the IA parameters.

While most of these methods have been developed to provide prescriptions for modelling IA in lensing  two-point statistics, some have been used to infuse galaxy alignments directly into cosmological simulations, as in \citet{Fluri2019, Tidalator2D, MICE_IA, Borg, Lanzieri}.  Having access to such IA-infused numerical simulations is crucial for a number of applications, including the possibility to validate theoretical models deep in the non-linear regime (or using non-linear galaxy bias model) \citep{IA_gal_bias}, to predict the impact of IA on non-Gaussian lensing probes  for which no model exists \citep{ZuercherXYZ, Tidalator2D}, to test IA mitigation techniques \citep{IA_selfCalibration}, or to learn about  the connection between large dark matter haloes and IA  \citep{vanAlfen}.  {\it (Any other important references?)}

This paper addresses several of the above-mentioned applications. After reviewing the modelling and measurement of cosmic shear data in Sec. \ref{sec:theory},  we describes our new flexible method to infuse different IA models in cosmological simulations (Sec. \ref{sec:IA_th}).  Specifically, we  infuse the NLA model, the $\delta$-NLA  and the TT model on the same underlying lensing simulation. In addition, we introduce the $\delta$-TT model (which takes into account the fact that galaxies trace dark matter even in the TT model), then proceed to couple the cosmic tidal fields with galaxies taken from halo occupation distributions (HOD) to investigate the impact of realistic non-linear galaxy bias on the IA signal. These new models are both physically motivated and challenging to describe theoretically as they require perturbation expansions beyond the second order.  The full numerical implementation is described in Sec. \ref{sec:sims} and validated against theoretical predictions at the level of  two-point shear correlation functions in Sec. \ref{sec:validation}.  Since our infusion method acts on shear galaxy catalogues and on convergence maps, we are in an ideal position to quantify the impact of multiple IA models on different non-Gaussian statistics, which we report in Sec. \ref{sec:HOWLS}, before concluding in Sec. \ref{sec:conclusion}.

 %----------
\section{Cosmic shear statistics}
\label{sec:theory}
%----------

Cosmic shear data can be analysed by a number of summary statistics, which all have their pros and cons. We first review the modelling and measurement methods of shear  two-point functions, then present other complementary cosmic shear statistics in the second part of this section.  
%-------------------------
\subsection{$\gamma$-2PCF}
\label{subsec:wl-th}

%---------------
\begin{figure}
\includegraphics[width=\columnwidth]{graphs/Nz}
\caption{Redshift distributions in our simulations, either taken from the Year-1 specifications for LSST (solid) or from a matched selection applied to an HOD galaxy catalogue (dashed, see Sec. \ref{subsec:HOD}).}
\label{fig:Nz}
\end{figure}
%--------------

 Shear two-point correlation functions ($\gamma$-2PCFs hereafter) can be predicted with percent-level precision and are therefore an ideal quantity to validate weak lensing simulations. In the Limber approximation\footnote{See \citet{Kilbinger17} for a comparison between the Limber approximation and the exact calculations.}, the lensing power spectrum $C^{ij}_{\ell}$ is obtained from an integral over the three-dimensional matter power spectrum $P_{\delta}(k,z)$:
  \begin{eqnarray}
C_{\ell}^{ij} = \int_0^{\chi_{\rm H}}  \frac{q^i(\chi) \,q^j(\chi)}{\chi^2} \, P_{\delta}\, \bigg(\frac{\ell+1/2}{\chi},z(\chi)\bigg) \ {\rm d}\chi,
\label{eq:C_ell}
\end{eqnarray}
where $\chi_{\rm H}$ is the co-moving distance to the horizon, $c$ is the speed of light, $H_0$ the Hubble parameter,  and the  lensing kernels $q^{i}$ are given by:
\begin{eqnarray}
q^i(\chi) = \frac{3}{2}\Omega_{\rm m} \, \bigg(\frac{H_0}{c} \bigg)^2 \frac{\chi}{a(\chi)} \int_{\chi}^{\chi_{\rm H}} n^i(\chi')\frac{{\rm d}z}{{\rm d} \chi'}\frac{\chi' - \chi}{\chi'}{\rm d}\chi' \, .
\label{eq:q_lensing}
\end{eqnarray}
In the previous expression, $n^i(z)$ refers to the redshift distribution  in tomographic bin `$i$', while $a(\chi)$ is the scale factor at comoving distance $\chi$ from the observer.  
 The matter power spectrum is computed from  {\sc Halofit} \citep{Takahashi2012} in this work, however  other public tools provide accurate predictions, including e.g. {\sc HMcode} \citep{HMCode2020}, the {\sc EuclidEmulator} \citep{EuclidEmulator}, the {\sc Bacco} emulator \citep{BACCOEmulator} or the {\sc MiraTitan} emulator \citep{miraTitan}. 
 Predictions for the $\gamma$-2PCFs are finally computed from Eq. (\ref{eq:C_ell}) as:
 \begin{eqnarray}
\xi_{\pm}^{ij}(\vartheta) = \frac{1}{2\pi}\int_0^{\infty} C_{\ell}^{ij} \, J_{0/4}(\ell \vartheta) \, \ell \, {\rm d}\ell,
\label{eq:xipm_th}
\end{eqnarray}
where $J_{0/4}(x)$ are Bessel functions of the first kind. In this paper, these calculations are carried out by the {\sc CosmoSIS} cosmological inference package \citep{cosmoSIS}. The $n^i(z)$ are taken from the LSST Science Requirement Document \citep{LSST-SRD} and given by:
 \begin{eqnarray}
n(z) = N_0 z^2 {\rm exp} \left[-\left(\frac{z}{z_0}\right)^\alpha \right]\, ,
\label{eq:nz}
\end{eqnarray}
with $(z_0, \alpha) = (0.13, 0.78)$ and $N_0$ normalised to provide a number density of $n_{\rm eff}$ = 3.0 gal arcmin$^{-2}$. This lower than the expected number density for the first data release ($n_{\rm eff}\sim10$ gal arcmin$^{-2}$), but is large enough to validate our methods, and makes the calculation more tractable. This global $n(z)$ is further split into five equi-populated tomographic bins, and smoothed with a Gaussian filter of width $\sigma = 0.05(1+z)$ to mimic the photometric selection process [{\it confirm with Niko}], shown by the solid lines in Fig. \ref{fig:Nz}.

The  weak lensing signal is  measured from the ellipticities $\epsilon_{1/2}$ of simulated or observed galaxies, which, in absence of systematics and secondary signals, are unbiased estimators  of the cosmic shear components $\gamma_{1/2}$. In particular, the $\gamma$-2PCF are estimated as:
 \begin{eqnarray}
 \widehat{\xi_{\pm}^{ij}}(\vartheta) = \frac{\sum_{a,b} w_a w_b \left[\epsilon_{a, \rm t}^i\epsilon_{b, \rm t}^j     \pm \epsilon_{a, \times}^i\epsilon_{b, \times}^j    \right]\Delta_{ab}(\vartheta)}{\sum_{a,b} w_a w_b} \, ,
 \label{eq:2PCF_estimator}
  \end{eqnarray}
where the sum is over all pairs of galaxies $a,b$ separated by an angular distance $\vartheta$ on the sky, respectively belonging to the tomographic bins $i$ and $j$. The tangential/cross components of their ellipticities are denoted as $\epsilon_{\rm t/\times}$, respectively. 
$\Delta_{ab}(\vartheta)$ is the binning operator, equal to unity if the angular separation between the two galaxies falls within the $\vartheta$-bin, and zero otherwise, while the shape weights $w_{a,b}$ are set to unity. In this paper, we construct lensing catalogues from numerical simulations (described in Sec. \ref{sec:sims}), from which we measure $ \widehat{\xi_{\pm}^{ij}}(\vartheta)$ with {\sc Treecorr} \citep{TreeCorr}. We set therein the {\sc bin\_slop} parameter to 0.05, then compute the correlations in 20 logarithmically-spaced angular bins with outer edges ranging from 0.5 to 475.5 arcmin.

 %-------------------------
\subsection{Non-Gaussian lensing statistics}
\label{subsec:beyond-2pt}

Non-Gaussian statistics have the potential to extract cosmological information stored in the complex phases of the density field, outperforming in that sense the $\gamma$-2PCF that can only access information encoded in the complex amplitudes. No optimal estimator has been identified to-date as being able to capture `all' information, but numerous studies all show a clear gain in constraining power, especially when used in combination with two-point functions \citep{HD21, DESY3_Zurcher, HSC_PeaksSims}. While some of these act on galaxy catalogues directly, many non-Gaussian statistics are measured on convergence or aperture mass maps, requiring the post-process the galaxy catalogues with a mass-reconstruction algorithm. To accommodate a variety of case we therefore provide both galaxy catalogues and convergence maps, the latter being  produced from the galaxies' observed ellipticities with  the standard Kaiser \& Squires inversion technique \citep{KaiserSquires}. These are spherical {\sc Healpix}\footnote{{\sc Healpix}: http://healpix.sf.net}  maps \citep{healpix} with $N_{\rm side}=4096$ generated over an octant,  and includes a baseline 2.0 arcmin smoothing beam to suppress numerical noise. Note that this smoothing scales is quite small; some probes are difficult to model theoretically in that regime, in which case the modelling needs to be inferred from simulations directly \citep[as in][]{HD21, DESY3_Zuercher, HSCY1_Peaks_sims}. We do not discuss statistics based on aperture mass maps in this paper for the sake of conciseness, but all our methods can be applied to those estimators as well.

Whereas tomographic $\gamma$-2PCFs data consists of auto-correlation and cross-redshift correlations, some non-Gaussian statistics can learn additional information from analysing catalogues or maps constructed from triplets, quadruplets or quintuplet of redshift bins \citep{Martinet21a}.  We do not fully exploit this possibility here and consider auto-tomographic bin only, nevertheless all of our methods and findings can be easily generalised to include this.

\subsection{Catalogue-based statistics}

A number of higher-order statistics work at the level of shear catalogue,  and here we consider the following:
\begin{enumerate}
\item {\it Three-point correlation functions (or $M_{\rm ap}^3(\vartheta)$)}:  [{\it Lucas/Laila, can you summarise the measurements and modelling, and provide references?}]
\item {\it Squeezed bispectrum}: [{\it Anik, can you summarise the measurements and modelling, and provide references?}]
\end{enumerate}

\subsection{Convergence-based statistics}

We further consider the following convergence map-based statistics:
\begin{enumerate}
\item {\it Convergence three-point correlation function}: [{\it Alejandro, can you summarise the measurements and modelling, and provide references?}]
\item {\it Peaks and Minima}: local maxima and minima in the {\sc Healpix} convergence maps are counted in bins of signal-to-noise ratios, $\nu$, computed from the global level of shape noise in the survey. These are amongst the most widely used non-Gaussian lensing statistics, as it is simple to measure yet highly sensitive to non-Gaussian structures. Although some theoretical models have been developed to describe the largest peaks \citep[see][]{Shan18, HSCY1_Peaks_th}, we report here the results on a much wider range of statistics, which can only be accurately modelled from numerical simulations.
\item {\it Lensing PDF}:  [{\it Cora/Lina, can you summarise the measurements and modelling, and provide references?}]
\item {\it Minkowski functionals}: [{\it Nisha, can you summarise the measurements and modelling, and provide references?}]
\end{enumerate}

 Each of these statistics probe different physical scales and non-linearities, causing their sensitivities to cosmological and systematics to vary. For example, a probe that is performing well at recovering cosmological information might be heavily affected by secondary signals such as IA and therefore deemed not robust. Having infused IA mocks at the field-level is therefore critical to answer such questions.

@@

Galaxies interact with the tidal forces caused by the large-scale structure  they live in, causing their intrinsic shapes to acquire correlated alignments that has nothing to do with the correlations caused by weak lensing. This therefore acts as a secondary signal that contaminates the shape correlation measurements carried out in a cosmic shear analysis. There is no consensus on the actual physical model that describes the IA signal, and even when adopting these, the free parameters they contain are only weakly constrained from the data \citep[see][for reviews]{IA_review1, IA_review2}.  %The most widely used model in the literature is the non-linear alignment model of \citet{NLA}, however it is now recognised that this is an effective model with limited precision, and the community is now progressively shifting towards more complex models. 
In this paper we consider two coupling models each applied on three galaxy bias cases, resulting in six different IA models, described in the following sub-sections. In all cases, our models couple the galaxy intrinsic shapes with the local over-density and projected tidal field, then prescribes an intrinsic ellipticity tensor $\gamma_{ij}^{\rm IA}$, from which intrinsic  ellipticities are extracted:
 \begin{eqnarray}
\epsilon_{1}^{\rm IA} = \gamma_{xx}^{\rm IA} - \gamma_{yy}^{\rm IA} \, , \hspace{1.5cm}\epsilon_{2}^{\rm IA} = 2 \gamma_{xy}^{\rm IA}\, .
\label{eq:tidal_th_deltaNLA}
\end{eqnarray}

@@

%The observed ellipticity of a galaxy ${\boldsymbol \epsilon}_{\rm obs} $ is a combination of its intrinsic shape ${\boldsymbol \epsilon}_{\rm int}$ and a cosmic shear signal ${\boldsymbol \gamma}$, the former of which can be further divided in a random component ${\boldsymbol \epsilon}^{\rm ran}$ and an alignment term ${\boldsymbol \epsilon}^{\rm IA}$.
The NLA model of \citet{NLA} is the most widely used IA model in the cosmic shear literature thus far. According to the NLA, IA are caused by a linear coupling between galaxy shapes and the non-linear large-scale tidal field at the galaxy position. The intrinsic ellipticities $\epsilon_{1,2}^{\rm NLA}$ are related to tidal field $s_{ij}$ by:
 \begin{eqnarray}
\epsilon_1^{\rm NLA} = - \frac{A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)} (s_{xx} - s_{yy}) \,\,\,\,,   \epsilon_2^{\rm NLA} = - \frac{2 A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)} s_{xy}\;,
\label{eq:tidal_th}
\end{eqnarray}
where $s_{ij} = \partial_{ij}\phi$ are the Cartesian components of the tidal tensor of the gravitational potential,  $D(z)$ is the linear growth factor, $\bar{\rho}(z)$ is the mean matter density at redshift $z$ and $\bar{C_1}=5\times 10^{-14} M_{\odot}^{-1} h^{-2} {\rm Mpc}^3$ is a constant calibrated in \citet{Brown2002}. The strength of the tidal coupling is modulated by the amplitude parameter $A_{\rm IA}$, which is the main NLA parameter constrained by current cosmic shear surveys. Note that this model can be augmented by redshift and luminosity dependences, however we do not use these here. Also note that the term `non-linear' in the model name can be  misleading, as it refers to the non-linear matter power spectrum $P(k)$ that is used in its calculations; the coupling between the intrinsic galaxy shapes and the tidal field is still linear and   Eq.( \ref{eq:tidal_th}) is used to assign intrinsic ellipticities to galaxies, given maps of the tidal field components $s_{xx}$, $s_{yy}$ and $s_{xy}$ (presented in  Sec. \ref {subsec:IA_infusion}).

In the context of  two-point functions, these intrinsic shapes contribute to an intrinsic-intrinsic ($II$) term and an intrinsic-shear coupling ($GI$) term \citep{Hirata2004}, both secondary signals to the true cosmic shear ($GG$) term, with the $GI$ typically dominating the IA sector in cross-tomographic measurements.  The $II$ and $GI$ terms  can be both computed from the matter power spectrum as: 
 \begin{eqnarray}
P_{II}(k,z) =  \left(\frac{A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)}\right)^2a^4(z) P_{\delta}(k,z)
\label{eq:Pk_II_th}
\end{eqnarray}
and
 \begin{eqnarray}
P_{GI}(k,z) = - \frac{A_{\rm IA}\bar{C_1}\bar{\rho}(z)}{D(z)}a^2(z) P_{\delta}(k,z) \, ,
\label{eq:Pk_GI_th}
\end{eqnarray}
which can then be past to the Limber integration (Eqs. \ref{eq:C_ell} and \ref{eq:xipm_th}) to compute  the secondary signals $\xi_{\pm}^{II}(\vartheta)$ and $\xi_{\pm}^{GI}(\vartheta)$.

@@

As shown in \citet{Blazek2019}, one-loop perturbative calculations include another term, by which galaxies acquire an intrinsic alignment via a coupling between their angular momentum and the tidal field, which can be re-expressed as a quadratic coupling between the tidal field and their shape. In this tidal torque model (TT), intrinsic ellipticities are given by:
\begin{eqnarray}
\gamma_{ij}^{\rm IA, TT} = C_2 \left[ \sum_{k=x,y,z} s_{ik} s_{kj} -\frac{1}{3} \delta_{ij} s^2 \right] \, ,
\label{eq:tidal_th_TT}
\end{eqnarray}
where 
\begin{eqnarray}
C_2 = \frac{5 A_2 \bar{C_1} \Omega_{\rm m} \rho_{\rm crit}}{D^2(z)} \,  = \left[ \frac{-5 A_2}{A_1 D(z)} \right] C_1.
\end{eqnarray}
Under the approximation that line-of-sight alignments are mostly suppressed from cosmic shear measurement due to the broad lensing kernels, 
we show in  Appendix \ref{app:2d_TT} that the terms inside the square bracket in Eq. (\ref{eq:tidal_th_TT}) leads to:
\begin{eqnarray}
\epsilon_1^{\rm TT} = C_2  \left[ s_{xx}^2 - s_{yy}^2\right] \, , \epsilon_2^{\rm TT} = 2 C_2 s_{xy}\left[s_{xx}+s_{yy}  \right]\, ,
\end{eqnarray}
both quadratic in the tidal field components. In this model, galaxies are also assumed to be randomly distributed on the sky; the $\delta$-weighted term requires third order perturbations.

@@

Finally, once all galaxies have been placed in the light-cone, we interpolate the shear and IA quantities at their exact location.

@@

JHD acknowledges support from an STFC Ernest Rutherford Fellowship (project reference ST/S004858/1). Ray-tracing computations were carried on the NERSC [acknowledgments] ... HACC [acknowledgements] ... {\it Outer Rim}[Acknowledgments]
\\

@@

%The SLICS numerical simulations can be found at http://slics.roe.ac.uk/, while the cosmo-SLICS can be made available upon request.
The construction of our tomographic $n^i(z)$ can be reproduced in {\tt github...CCL\_SRD\_Niko}.
